---
title: Modeling Protection from COVID-19 Based on Vaccine Supply and Administration
  Rates
author: "Kieran Allsop, James C. Capretta, and Scott Ganz"
date: "2/24/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(readr)
library(stringr)
library(purrr)
library(ggthemes)
library(ggplot2)
library(scales)
```

## Introduction

This Markdown file replicates the stock and flow model used by Kieran Allsop, James C. Capretta, and Scott Ganz in the AEI Economic Perspectives publication [Modeling Protection from COVID-19 Based on Vaccine Supply and Administration](). The code was originally run on R.4.0.1. Updated versions of R may have been released since this replication file was written.

Note that data for the model updates daily and as such estimations may change over time. The model provided is meant to provide a snapshot of the current situation. Current values are correct as of 02/16/2021, the date from which data was taken for the report.

You will need to install the following packages into R to run this code:

* dplyr
* tidyr
* lubridate
* readr
* ggplot2
* ggthemes
* stringr
* scales
* purrr

## Initialize Objects and Set Parameters

First, initialize the objects needed to store data, and set parameters for population and time for your model.

```{r initialize}
## -- INITIALIZE OBJECTS -- ##
stock <- list()
deliveries <- list()
schedule <- list()
outcomes <- list()
vaccinations <- list()

## -- SET PARAMETERS -- ##
weeks <- 33
start_date <- "2021-02-15"
US_population <- 328239523

```

## Set Initializing Data Points

Data on current vaccination figures is provided by the [CDC](https://covid.cdc.gov/covid-data-tracker/#vaccinations). Shipment schedules for [Moderna](https://data.cdc.gov/Vaccinations/COVID-19-Vaccine-Distribution-Allocations-by-Juris/b7pe-5nws) and [Pfizer-BioNTech](https://data.cdc.gov/Vaccinations/COVID-19-Vaccine-Distribution-Allocations-by-Juris/saz5-9hgg) are also provided by the CDC. 

```{r initial_data}
# Read in Moderna shipment schedule
moderna_shipping <-
  read_csv(
    "../data/COVID-19_Vaccine_Distribution_Allocations_by_Jurisdiction_-_Moderna_0216.csv"
  ) %>%
  filter(Jurisdiction == "Total") %>%
  select(-1,-2) %>%
  mutate_all( ~ str_replace_all(., ",", "") %>% as.numeric)

# Read in Pfizer shipment schedule
pfizer_shipping <-
  read_csv(
    "../data/COVID-19_Vaccine_Distribution_Allocations_by_Jurisdiction_-_Pfizer_0216.csv"
  ) %>%
  filter(Jurisdiction == "Total") %>%
  select(-1,-2) %>%
  mutate_all( ~ str_replace_all(., ",", "") %>% as.numeric)

# Input column numbers of weeks still due to receive second doses 
# (3wks for Pfizer, 4 wks for Moderna)
pfizer_second <-
  pfizer_shipping[, c(14, 16, 18)] %>% unlist(use.names = F)
moderna_second <-
  moderna_shipping[, c(10, 12, 14, 16)] %>% unlist(use.names = F)

# Add they scheduled second doses to a new "schedule" object
schedule[[1]] <- list()
schedule[[1]][["pfizer"]] <- tibble(week = c(1, 2, 3),
                                    people = pfizer_shipping[, c(14, 16, 18)] %>% unlist(use.names = F))
schedule[[1]][["moderna"]] <- tibble(week = c(1, 2, 3, 4),
                                     people = moderna_shipping[, c(10, 12, 14, 16)] %>% unlist(use.names = F))

# Input initialization values from the CDC
# Up to date as of 02/16/21
doses_delivered0 <- 71657975
doses_administered0 <- 55220364
doses_administered_pfizer0 <- 28374410
doses_administered_moderna0 <- 26738383
doses_administered_unknown0 <- 107751
people_fully_vaccinated0 <- 15015434
people_partially_vaccinated0 <- 39670551 - people_fully_vaccinated0
doses_stock0 <- doses_delivered0 - doses_administered0
doses_scheduled_pfizer0 <- sum(schedule[[1]][["pfizer"]]$people)
doses_scheduled_moderna0 <- sum(schedule[[1]][["moderna"]]$people)

# Calculate doses delivered and doses in stock for each of the respectives vaccine candidates
doses_delivered_pfizer0 <-
  doses_delivered0 * doses_administered_pfizer0 / (doses_administered_moderna0 + doses_administered_pfizer0)
doses_delivered_moderna0 <-
  doses_delivered0 * doses_administered_moderna0 / (doses_administered_moderna0 + doses_administered_pfizer0)
doses_stock_pfizer0 <-
  doses_scheduled_pfizer0 + (doses_stock0 - doses_scheduled_moderna0 - doses_scheduled_pfizer0) * doses_administered_moderna0 /
  (doses_administered_moderna0 + doses_administered_pfizer0)
doses_stock_moderna0 <-
  doses_scheduled_moderna0 + (doses_stock0 - doses_scheduled_moderna0 - doses_scheduled_pfizer0) * doses_administered_pfizer0 /
  (doses_administered_moderna0 + doses_administered_pfizer0)

```

## Set up Future Delivery Schedules

Pfizer-BioNTech, Moderna, and Johnson & Johnson have all announced deadlines for deliveries of certain amounts of doses of their vaccines. Set those here.

```{r delivery_schedule}
## -- FLOW -- ##

# When deliveries are spread over time, assume that the roll out is evenly 
# distributed unless otherwise noted

p1 <- 7     # Present - Mar 31 (Q1)
p2 <- 8     # Mar 31 - May 31
p3 <- 9     # May 31 - July 31
p1jj <- 2   # Mar 15 - Mar 31
p2jj <- 4   # Mar 31 - Apr 30
p3jj <- 9   # Apr 30 - Jun 30
Endjj <- weeks - p1jj - p2jj - p3jj - 6
End <- weeks - p1 - p2 - p3

## -- PFIZER -- ##
# Current schedule: 120MM in Q1, 200MM by May 31, 300MM by Jul 31
doses_weekly_pfizer_p1 <- (120 * 10 ^ 6 - doses_delivered_pfizer0) / p1
doses_weekly_pfizer_p2 <- (80 * 10 ^ 6) / p2
doses_weekly_pfizer_p3 <- (100 * 10 ^ 6) / p3
deliveries[["pfizer"]] <- tibble(week = 1:weeks,
                                 doses = c(
                                   rep(doses_weekly_pfizer_p1, p1),
                                   rep(doses_weekly_pfizer_p2, p2),
                                   rep(doses_weekly_pfizer_p3, p3),
                                   rep(0, End)
                                 ))

## -- MODERNA -- ##
# Current schedule: 100MM in Q1, 200MM by May 31, 300MM by Jul 31
doses_weekly_moderna_p1 <- (100 * 10 ^ 6 - doses_delivered_moderna0) / p1
doses_weekly_moderna_p2 <- 100 * 10 ^ 6 / p2
doses_weekly_moderna_p3 <- 100 * 10 ^ 6 / p3
deliveries[["moderna"]] <- tibble(week = 1:weeks,
                                  doses = c(
                                    rep(doses_weekly_moderna_p1, p1),
                                    rep(doses_weekly_moderna_p2, p2),
                                    rep(doses_weekly_moderna_p3, p3),
                                    rep(0, End)
                                  ))

## -- WITHOUT JOHNSON & JOHNSON -- ##
# Current schedule: 7MM by Mar 31, 20MM by Apr 30, 100MM by Jun 30
doses_weekly_jj_p1 <- 0 / p1jj
doses_weekly_jj_p2 <- 0 / p2jj
doses_weekly_jj_p3 <- 0 / p3jj
deliveries[["jj"]] <- tibble(week = 1:weeks,
                             doses = c(rep(0, 6),
                                       rep(doses_weekly_jj_p1, p1jj),
                                       rep(doses_weekly_jj_p2, p2jj),
                                       rep(doses_weekly_jj_p3, p3jj),
                                       rep(0, Endjj)
                                       ))
```

## Set Up Capacity Constraints

Different scenarios are built on different capacity constraints. The current rate of increase in daily vaccinations is calculated using data from [Our World in Data](https://github.com/owid/covid-19-data/tree/master/public/data/vaccinations). Current capacity is taken from [Bloomberg](https://www.bloomberg.com/graphics/covid-vaccine-tracker-global-distribution/).

```{r capacity_constraints}
## -- CAPACITY CONSTRAINTS -- ##

# Current capacity: up to date as of 02/16/21
capacity0 <- 1.67 * 10^6

# Calculate growth rate
owid_url <-
  "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/us_state_vaccinations.csv"
owid_data <-
  read_csv(
    url(
      owid_url <-
        "https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/us_state_vaccinations.csv"
    )
  ) %>%
  filter(location == "United States",!is.na(daily_vaccinations)) %>%
  select(date, total_vaccinations, daily_vaccinations) %>%
  filter(date >= ymd("2021-01-01")) %>%
  filter(date <= ymd("2021-02-16")) %>%
  mutate(day = 1:n())

rate <- coef(lm(daily_vaccinations ~ day, data = owid_data))[2]
rm(owid_url, owid_data)

## -- DIFF. MAX CAPCITY CEILINGS -- ##

# Max capacity: 3 million/day (Base Case)
capacity3 <- tibble(week = 1:weeks) %>%
  mutate(
    daily_capacity = capacity0 + (7 * rate) * week,
    daily_capacity = ifelse(daily_capacity > 3 * 10 ^ 6, 3 * 10 ^
                              6, daily_capacity),
    weekly_capacity = daily_capacity * 7
  )

# Max capacity: 3.5 million/day (High Case)
capacity35 <- tibble(week = 1:weeks) %>%
  mutate(
    daily_capacity = capacity0 + (7 * rate) * week,
    daily_capacity = ifelse(daily_capacity > 3.5 * 10 ^ 6, 3.5 * 10 ^
                              6, daily_capacity),
    weekly_capacity = daily_capacity * 7
  )

# Max capacity: 2.5 million/day (Low Case)
capacity25 <- tibble(week = 1:weeks) %>%
  mutate(
    daily_capacity = capacity0 + (7 * rate) * week,
    daily_capacity = ifelse(daily_capacity > 2.5 * 10 ^ 6, 2.5 * 10 ^
                              6, daily_capacity),
    weekly_capacity = daily_capacity * 7
  )

# Max capacity: current capacity
capacitynow <- tibble(week=1:weeks) %>%
  mutate(daily_capacity = capacity0,
         weekly_capacity = daily_capacity * 7)

# J&J Max capacity: 1 million/day (Only for J&J)
capacityjj <- tibble(week = 1:weeks) %>%
  mutate(daily_capacity = 1 * 10^6,
         weekly_capacity = daily_capacity * 7)
```

## Final Objects to Initialize

Finally, initialize a stock object to store data on the total stock and a population object to keep track of how many people have had a vaccine and how many doses they've had.

```{r initialize2}
## -- STOCK -- ##
stock <- list()
stock[[1]] <- list()
stock[[1]][["moderna"]] <- doses_stock_moderna0
stock[[1]][["pfizer"]] <- doses_stock_pfizer0
stock[[1]][["jj"]] <- 0

## -- POPULATION CONSTRAINT -- ##
people <- list()
people[[1]] <- list()
people[[1]][["partial"]] <- people_partially_vaccinated0
people[[1]][["completed"]] <- people_fully_vaccinated0
people[[1]][["unprotected"]] <- US_population - people[[1]][["completed"]] - people[[1]][["partial"]]
```

## Create Functions for the Model

These functions will be used to update objects as the model processes each week.

* **init_vaccinations**: This function compiles all data for the week in the object 'vaccinations' including stock at the beginning of the week, deliveries at the start of that week, capacity for that week, and the schedule of people requiring second doses that week. It also creates an 'administration' list within the 'vaccination' object.
* **give_vaccinations**: Based on capacity, deliveries, and the schedule of people needing second doses, this function decides how many vaccinations to give out and of what type.
* **update_schedule**: This function updates the number of people needing second shots of either of the two mRNA vaccines in the coming weeks.
* **check_pop_limits**: This function checks to make sure we have not vaccinated more people than there is in the population and corrects administration numbers if we have.
* **update_stock**: This function updates the number of doses of each type of vaccine in the current stock.
* **update_people**: This function updates the 'people' object that tells us how many people are completely vaccinated, how many are partially vaccinated, and how many people are still unprotected.
* **update_outcomes**: This function collects the relevant information for the week and stores it in the outcomes object
* **create_outcomes_table**: This function creates the table in which all the data from each week will appear and the tables that we will make our graphs from. 

```{r functions}
## -- INITIALIZE A WEEK -- ##
init_vaccinations = function(stock, deliveries, capacity, capacityjj, schedule, wk){
  
  vaccinations <- list()
  vaccinations[["stock"]] <- list()
  vaccinations[["stock"]][["moderna"]] <- stock[[wk]]$moderna
  vaccinations[["stock"]][["pfizer"]] <- stock[[wk]]$pfizer
  vaccinations[["stock"]][["jj"]] <- stock[[wk]]$jj
  
  vaccinations[["deliveries"]] <- list()
  vaccinations[["deliveries"]][["moderna"]] <-
    deliveries[["moderna"]] %>% filter(week == wk) %>% pull(doses)
  vaccinations[["deliveries"]][["pfizer"]] <-
    deliveries[["pfizer"]] %>% filter(week == wk) %>% pull(doses)
  vaccinations[["deliveries"]][["jj"]] <-
    deliveries[["jj"]] %>% filter(week == wk) %>% pull(doses)
  
  vaccinations[["capacity"]] <-
    capacity %>% filter(week == wk) %>% pull(weekly_capacity)
  vaccinations[["capacityjj"]] <-
    capacityjj %>% filter(week == wk) %>% pull(weekly_capacity)
  
  vaccinations[["second_doses"]] <- list()
  vaccinations[["second_doses"]][["moderna"]] <-
    schedule[[wk]]$moderna %>% filter(week == wk) %>% pull(people)
  vaccinations[["second_doses"]][["pfizer"]] <-
    schedule[[wk]]$pfizer %>% filter(week == wk) %>% pull(people)
  
  vaccinations[["administration"]] <- list()
  
  return(vaccinations)
}

## -- GIVE MODERNA AND PFIZER VACCINATIONS -- ##
give_vaccinations = function(vaccinations, capacity, capacityjj, people, stock, deliveries, schedule, wk){
  
  administration <- list()
  
  # case 1: capacity is not sufficient to cover scheduled doses - give vaccines
  # in proportion to people waiting
  if (vaccinations[[wk]][["capacity"]] <=
      vaccinations[[wk]][["second_doses"]][["moderna"]] +
      vaccinations[[wk]][["second_doses"]][["pfizer"]]) {

      administration[["pfizer_second"]] <-
        capacity * vaccinations[[wk]][["second_doses"]][["pfizer"]] /
        (vaccinations[[wk]][["second_doses"]][["pfizer"]] +
        vaccinations[[wk]][["second_doses"]][["moderna"]])
      administration[["moderna_second"]] <-
        capacity * vaccinations[[wk]][["second_doses"]][["moderna"]] /
        (vaccinations[[wk]][["second_doses"]][["pfizer"]] +
        vaccinations[[wk]][["second_doses"]][["moderna"]])
      administration[["moderna_first"]] <- 0
      administration[["pfizer_first"]] <- 0
      
    # J&J
    max_jj <- (stock[[wk]][["jj"]] +
                 deliveries[["jj"]] %>% filter(week == wk) %>% pull(doses))
    
    if (vaccinations[[wk]][["capacityjj"]] <= max_jj) {
      administration[["jj"]] <- vaccinations[[wk]][["capacityjj"]]
    }else{
      administration[["jj"]] <- max_jj
    }
    
  # case 2: if no one needs a first dose of the vaccine
  } else if (people[[wk]]$unprotected == 0) {
      administration[["pfizer_second"]] <-
        vaccinations[[wk]][["second_doses"]][["pfizer"]]
      administration[["moderna_second"]] <-
        vaccinations[[wk]][["second_doses"]][["moderna"]]
      administration[["pfizer_first"]] <- 0
      administration[["moderna_first"]] <- 0
      administration[["jj"]] <- 0
  } else {
      max_moderna_first <- floor((1/2) * (stock[[wk]][["moderna"]] + 
                                          deliveries[["moderna"]] %>% 
                                          filter(week == wk) %>% pull(doses) -
                                          sum((schedule[[wk]][["moderna"]] %>% pull(people)))))
    
      max_pfizer_first <- floor((1/2) * (stock[[wk]][["pfizer"]] +
                                          deliveries[["pfizer"]] %>% 
                                          filter(week == wk) %>% pull(doses) -
                                         sum((schedule[[wk]][["pfizer"]] %>% pull(people)))))
      # case 3: if capacity is sufficient to cover scheduled doses but not the 
      # maximum amount of first doses
      if (vaccinations[[wk]][["capacity"]] <=
        vaccinations[[wk]][["second_doses"]][["moderna"]] +
        vaccinations[[wk]][["second_doses"]][["pfizer"]] +
        max_moderna_first +
        max_pfizer_first) {
          administration[["pfizer_second"]] <-
            vaccinations[[wk]][["second_doses"]][["pfizer"]]
          administration[["moderna_second"]] <-
            vaccinations[[wk]][["second_doses"]][["moderna"]]
          administration[["pfizer_first"]] <-
            (vaccinations[[wk]][["capacity"]] -
            administration[["pfizer_second"]] -
            administration[["moderna_second"]]) *
            max_pfizer_first / (max_pfizer_first + max_moderna_first)
          administration[["moderna_first"]] <-
            (vaccinations[[wk]][["capacity"]] -
            administration[["pfizer_second"]] -
            administration[["moderna_second"]]) *
            max_moderna_first / (max_pfizer_first + max_moderna_first)
      # case 4: if capacity is sufficient to cover scheduled doses and the
      # maximum amount of fist doses
      } else {
          administration[["pfizer_second"]] <-
            vaccinations[[wk]][["second_doses"]][["pfizer"]]
          administration[["moderna_second"]] <-
            vaccinations[[wk]][["second_doses"]][["moderna"]]
          administration[["pfizer_first"]] <-
            max_pfizer_first
          administration[["moderna_first"]] <-
            max_moderna_first
      }
      
      # J&J
      max_jj <- (stock[[wk]][["jj"]] +
                 deliveries[["jj"]] %>% filter(week == wk) %>% pull(doses))
    
      if (vaccinations[[wk]][["capacityjj"]] <= max_jj) {
        administration[["jj"]] <- vaccinations[[wk]][["capacityjj"]]
      }else{
        administration[["jj"]] <- max_jj
      }
  }
  
  return(administration)
}

## -- UPDATE SCHEDULE -- ##
update_schedule = function(schedule, vaccinations, wk){
  
  new_schedule <- list()
  
  new_schedule[["pfizer"]] <-
    tibble(
      week = (wk + 1):(wk + 3),
      people = c(
        schedule[[wk]][["pfizer"]] %>% filter(week == wk + 1) %>% pull(people),
        schedule[[wk]][["pfizer"]] %>% filter(week == wk + 2) %>% pull(people),
        vaccinations[[wk]][["administration"]][["pfizer_first"]]
      )
    )
    
  new_schedule[["moderna"]] <-
    tibble(
      week = (wk + 1):(wk + 4),
      people = c(
        schedule[[wk]][["moderna"]] %>% filter(week == wk + 1) %>% pull(people),
        schedule[[wk]][["moderna"]] %>% filter(week == wk + 2) %>% pull(people),
        schedule[[wk]][["moderna"]] %>% filter(week == wk + 3) %>% pull(people),
        vaccinations[[wk]][["administration"]][["moderna_first"]]
      )
    )
  
  return(new_schedule)
}

## -- CHECK POPULATION LIMITS -- ##
check_pop_limits = function(vaccinations, people, wk){
  
  new_administration <- list()
  
  if (vaccinations[[wk]][["administration"]][["jj"]] >= abs(people[[wk + 1]][["unprotected"]])) {
    new_administration[["jj"]] <-  
      vaccinations[[wk]][["administration"]][["jj"]] - abs(people[[wk + 1]][["unprotected"]])
    new_administration[["pfizer_first"]] <- vaccinations[[wk]][["administration"]][["pfizer_first"]]
    new_administration[["pfizer_second"]] <- vaccinations[[wk]][["administration"]][["pfizer_second"]]
    new_administration[["moderna_first"]] <- vaccinations[[wk]][["administration"]][["moderna_first"]]
    new_administration[["moderna_second"]] <- vaccinations[[wk]][["administration"]][["moderna_second"]]
  } else if ((vaccinations[[wk]][["administration"]][["jj"]] +
              vaccinations[[wk]][["administration"]][["pfizer_first"]]) >= 
              abs(people[[wk + 1]][["unprotected"]])) {
    new_administration[["pfizer_first"]] <-
      vaccinations[[wk]][["administration"]][["pfizer_first"]] - 
      (abs(people[[wk + 1]][["unprotected"]]) - vaccinations[[wk]][["administration"]][["jj"]])
    new_administration[["jj"]] <- 0
    new_administration[["pfizer_second"]] <- vaccinations[[wk]][["administration"]][["pfizer_second"]]
    new_administration[["moderna_first"]] <- vaccinations[[wk]][["administration"]][["moderna_first"]]
    new_administration[["moderna_second"]] <- vaccinations[[wk]][["administration"]][["moderna_second"]]
  } else {
    new_administration[["moderna_first"]] <-
      vaccinations[[wk]][["administration"]][["moderna_first"]] - 
      (abs(people[[wk + 1]][["unprotected"]]) - 
      vaccinations[[wk]][["administration"]][["jj"]] -
      vaccinations[[wk]][["administration"]][["pfizer_first"]])
    new_administration[["jj"]]<- 0
    new_administration[["pfizer_first"]] <- 0
    new_administration[["pfizer_second"]] <- vaccinations[[wk]][["administration"]][["pfizer_second"]]
    new_administration[["moderna_second"]] <- vaccinations[[wk]][["administration"]][["moderna_second"]]
  }
  
  return(new_administration)
}

## -- UPDATE STOCK -- ##
update_stock = function(stock, vaccinations, deliveries, wk){
  
  new_stock <- list()
  
  new_stock[["moderna"]] <- stock[[wk]][["moderna"]] -
    vaccinations[[wk]][["administration"]][["moderna_second"]] -
    vaccinations[[wk]][["administration"]][["moderna_first"]] +
    (deliveries$moderna %>% filter(week == wk) %>% pull(doses))
  
  new_stock[["pfizer"]] <- stock[[wk]][["pfizer"]] -
    vaccinations[[wk]][["administration"]][["pfizer_second"]] -
    vaccinations[[wk]][["administration"]][["pfizer_first"]] +
    (deliveries$pfizer %>% filter(week == wk) %>% pull(doses))
  
  new_stock[["jj"]] <- stock[[wk]][["jj"]] -
    vaccinations[[wk]][["administration"]][["jj"]] +
    (deliveries$jj %>% filter(week == wk) %>% pull(doses))
  
  return(new_stock)
}

## -- UPDATE PEOPLE -- ##
update_people = function(people, schedule, vaccinations, US_population, wk){
  
  new_people <- list()
  
  new_people[["partial"]] <- 
    sum(schedule[[wk + 1]][["moderna"]]$people) + sum(schedule[[wk + 1]][["pfizer"]]$people)
  
  new_people[["completed"]] <- people[[wk]][["completed"]] + 
    vaccinations[[wk]][["administration"]][["pfizer_second"]] +
    vaccinations[[wk]][["administration"]][["moderna_second"]] +
    vaccinations[[wk]][["administration"]][["jj"]]
  
  new_people[["unprotected"]] <- 
    US_population - new_people[["completed"]] - new_people[["partial"]]
  
  return(new_people)
}

## -- UPDATE OUTCOMES -- ##
update_outcomes = function(outcomes, vaccinations, wk){
  
  new_outcomes <- list()
  
  new_outcomes[["doses_administered"]] <-
    unlist(vaccinations[[wk]][["administration"]]) %>% sum
  new_outcomes[["capacity"]] <- vaccinations[[wk]][["capacity"]]
  new_outcomes[["capacity_utilization"]] <-
    new_outcomes[["doses_administered"]] / vaccinations[[wk]][["capacity"]]
  new_outcomes[["stock_remaining"]] <-
    unlist(stock[[wk + 1]]) %>% sum
  new_outcomes[["vaccinations_completed"]] <-
    vaccinations[[wk]][["administration"]][["pfizer_second"]] +
    vaccinations[[wk]][["administration"]][["moderna_second"]] +
    vaccinations[[wk]][["administration"]][["jj"]]
  
  return(new_outcomes)
}

## -- CREATE OUTCOMES TABLE -- ##
create_outcomes_table = function(outcomes, stock, deliveries, start_date){
  data_table <- tibble(wk = 1:weeks,
                      date = ymd(start_date) + weeks(0:(weeks - 1)),
                      doses_administered_wk = map(outcomes, "doses_administered") %>%
                        unlist(),
                      vaccinations_completed_wk = map(outcomes, "vaccinations_completed") %>% 
                        unlist(),
                      capacity_wk = map(outcomes, "capacity") %>% unlist,
                      stock_wk = (map(stock, "moderna") %>% unlist +
                                  map(stock, "pfizer") %>% unlist +
                                  map(stock, "jj") %>% unlist)[-1],
                      stock_pfizer_wk = (map(stock, "pfizer") %>% unlist)[-1],
                      stock_moderna_wk = (map(stock, "moderna") %>% unlist)[-1],
                      stock_jj_wk = (map(stock, "jj") %>% unlist)[-1],
                      deliveries_pfizer_wk = deliveries$pfizer$doses,
                      deliveries_moderna_wk = deliveries$moderna$doses,
                      deliveries_jj_wk = deliveries$jj$doses,
                      doses_administered_tot = doses_administered0 + cumsum(doses_administered_wk),
                      vaccinations_completed_tot = people_fully_vaccinated0 + cumsum(vaccinations_completed_wk))
  return(data_table)
}
```

## Run Model Excluding Johnson & Johnson Candidate

Run the loop for the number of weeks given at the start with the functions just defined.

```{r withoutjj}
## -- RUN 1 - BASE CASE -- ##
capacity <- capacity3

for (wk in 1:weeks) {
  vaccinations[[wk]] <- init_vaccinations(stock, deliveries, capacity, capacityjj, schedule, wk)
  vaccinations[[wk]][["administration"]] <- 
    give_vaccinations(vaccinations, capacity, capacityjj, people, stock, deliveries, schedule, wk)
  schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
  stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
  people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
  outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  if (people[[wk + 1]]$unprotected < 0) {
    vaccinations[[wk]][["administration"]] <- check_pop_limits(vaccinations, people, wk)
    schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
    stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
    people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
    outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  }
}
outcomes_table_nojj_base <- create_outcomes_table(outcomes, stock, deliveries, start_date)

## -- RUN 2 - LOW CASE -- ##
capacity <- capacity25

for (wk in 1:weeks) {
  vaccinations[[wk]] <- init_vaccinations(stock, deliveries, capacity, capacityjj, schedule, wk)
  vaccinations[[wk]][["administration"]] <- 
    give_vaccinations(vaccinations, capacity, capacityjj, people, stock, deliveries, schedule, wk)
  schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
  stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
  people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
  outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  if (people[[wk + 1]]$unprotected < 0) {
    vaccinations[[wk]][["administration"]] <- check_pop_limits(vaccinations, people, wk)
    schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
    stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
    people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
    outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  }
}
outcomes_table_nojj_low <- create_outcomes_table(outcomes, stock, deliveries, start_date)

## -- RUN 3 - HIGH CASE -- ##
capacity <- capacity35

for (wk in 1:weeks) {
  vaccinations[[wk]] <- init_vaccinations(stock, deliveries, capacity, capacityjj, schedule, wk)
  vaccinations[[wk]][["administration"]] <- 
    give_vaccinations(vaccinations, capacity, capacityjj, people, stock, deliveries, schedule, wk)
  schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
  stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
  people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
  outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  if (people[[wk + 1]]$unprotected < 0) {
    vaccinations[[wk]][["administration"]] <- check_pop_limits(vaccinations, people, wk)
    schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
    stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
    people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
    outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  }
}
outcomes_table_nojj_high <- create_outcomes_table(outcomes, stock, deliveries, start_date)

```

## Change Model to Include Johnson & Johnson Candidate

Change the delivery schedule for the Johnson & Johnson candidate so that it is not zero and then re-run the model for all three cases.

```{r withjj}
## -- WITH JOHNSON & JOHNSON SCHEDULE -- ##
# Current schedule: 7MM by Mar 31, 20MM by Apr 30, 100MM by Jun 30
doses_weekly_jj_p1 <- 7 * 10 ^ 6 / p1jj
doses_weekly_jj_p2 <- 13 * 10 ^ 6 / p2jj
doses_weekly_jj_p3 <- 80 * 10 ^ 6 / p3jj
deliveries[["jj"]] <- tibble(week = 1:weeks,
                             doses = c(rep(0, 6),
                                       rep(doses_weekly_jj_p1, p1jj),
                                       rep(doses_weekly_jj_p2, p2jj),
                                       rep(doses_weekly_jj_p3, p3jj),
                                       rep(0, Endjj)
                                       ))

## -- RUN 1 - BASE CASE -- ##
capacity <- capacity3
# wk <- 21
for (wk in 1:weeks) {
  vaccinations[[wk]] <- init_vaccinations(stock, deliveries, capacity, capacityjj, schedule, wk)
  vaccinations[[wk]][["administration"]] <- 
    give_vaccinations(vaccinations, capacity, capacityjj, people, stock, deliveries, schedule, wk)
  schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
  stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
  people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
  outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  if (people[[wk + 1]]$unprotected < 0) {
    vaccinations[[wk]][["administration"]] <- check_pop_limits(vaccinations, people, wk)
    schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
    stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
    people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
    outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  }
}
outcomes_table_withjj_base <- create_outcomes_table(outcomes, stock, deliveries, start_date)

## -- RUN 2 - LOW CASE -- ##
capacity <- capacity25

for (wk in 1:weeks) {
  vaccinations[[wk]] <- init_vaccinations(stock, deliveries, capacity, capacityjj, schedule, wk)
  vaccinations[[wk]][["administration"]] <- 
    give_vaccinations(vaccinations, capacity, capacityjj, people, stock, deliveries, schedule, wk)
  schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
  stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
  people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
  outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  if (people[[wk + 1]]$unprotected < 0) {
    vaccinations[[wk]][["administration"]] <- check_pop_limits(vaccinations, people, wk)
    schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
    stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
    people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
    outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  }
}
outcomes_table_withjj_low <- create_outcomes_table(outcomes, stock, deliveries, start_date)

## -- RUN 3 - HIGH CASE -- ##
capacity <- capacity35

for (wk in 1:weeks) {
  vaccinations[[wk]] <- init_vaccinations(stock, deliveries, capacity, capacityjj, schedule, wk)
  vaccinations[[wk]][["administration"]] <- 
    give_vaccinations(vaccinations, capacity, capacityjj, people, stock, deliveries, schedule, wk)
  schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
  stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
  people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
  outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  if (people[[wk + 1]]$unprotected < 0) {
    vaccinations[[wk]][["administration"]] <- check_pop_limits(vaccinations, people, wk)
    schedule[[wk + 1]] <- update_schedule(schedule, vaccinations, wk)
    stock[[wk + 1]] <- update_stock(stock, vaccinations, deliveries, wk)
    people[[wk + 1]] <- update_people(people, schedule, vaccinations, US_population, wk)
    outcomes[[wk]] <- update_outcomes(outcomes, vaccinations, wk)
  }
}
outcomes_table_withjj_high <- create_outcomes_table(outcomes, stock, deliveries, start_date)
```

## Create Figures

Create Figures 3, 4, 5, and 6 from the report here.

```{r figures}
## -- FIGURE 3 - STOCK OVER TIME - NO J&J -- ##
# Create stock over time dataframe to create graph from
stock_nojj_base <- outcomes_table_nojj_base[, c("date", "stock_wk")] %>%
  rename(base = stock_wk)
stock_nojj_low <- outcomes_table_nojj_low[, c("date", "stock_wk")] %>%
  rename(low = stock_wk)
stock_nojj_high <- outcomes_table_nojj_high[, c("date", "stock_wk")] %>%
  rename(high = stock_wk)
stock_over_time_nojj <- merge(stock_nojj_low, stock_nojj_base)
stock_over_time_nojj <- merge(stock_over_time_nojj, stock_nojj_high)
stock_over_time_nojj <- stock_over_time_nojj %>%
  rename('Max 2.5 million/day' = low, 'Max 3 million/day' = base,
         'Max 3.5 million/day' = high) %>%
  select(date, "Max 2.5 million/day", "Max 3 million/day", "Max 3.5 million/day") %>%
  gather(key = "Scenario", value = "value", -date)

# Create graph
stock_graph_nojj <- ggplot(stock_over_time_nojj, aes(x=date, y=value)) +
  geom_line(aes(color = Scenario)) + 
  geom_point(aes(color=Scenario)) +
  scale_color_manual(values = c("darkred", "orange", "darkgreen")) +
  xlab("Date") + ylab("Total Stock") + 
  theme_few() + scale_y_continuous(labels=comma) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme(panel.grid.major.y = element_line(colour="lightgrey", size = .25)) +
  ggtitle("Total Stock of all Vaccines - Without J&J")
stock_graph_nojj

## -- FIGURE 4 - PCT POPULATION VACCINATION - NO J&J -- ##
# Create percent vaccinated variables
outcomes_table_nojj_base <- outcomes_table_nojj_base %>%
  mutate(base = vaccinations_completed_tot/US_population)
outcomes_table_nojj_low <- outcomes_table_nojj_low %>%
  mutate(low = vaccinations_completed_tot/US_population)
outcomes_table_nojj_high <- outcomes_table_nojj_high %>%
  mutate(high = vaccinations_completed_tot/US_population)

# Create a new dataframe with percent vaccinated for graph
pct_vacc_table_nojj <- outcomes_table_nojj_base[,c("date", "base")]
pct_vacc_table_nojj <- cbind(pct_vacc_table_nojj, outcomes_table_nojj_low$low)
pct_vacc_table_nojj <- cbind(pct_vacc_table_nojj, outcomes_table_nojj_high$high)
pct_vacc_table_nojj <- pct_vacc_table_nojj %>%
  rename('Max 2.5 million/day' = 'outcomes_table_nojj_low$low', 
         'Max 3 million/day' = base,
         'Max 3.5 million/day' = 'outcomes_table_nojj_high$high') %>%
  select(date, "Max 2.5 million/day", "Max 3 million/day", "Max 3.5 million/day") %>%
  gather(key = "Scenario", value = "value", -date)

# Create graph
population_vaccinated_nojj <- ggplot(pct_vacc_table_nojj, aes(x=date, y=value)) +
  geom_line(aes(color = Scenario)) + 
  geom_point(aes(color=Scenario)) +
  scale_color_manual(values = c("darkred", "orange", "darkgreen")) +
  geom_hline(yintercept = 0.5, color="lightblue", size = 1) +
  geom_hline(yintercept = 0.75, color="blue", size = 1) +
  geom_hline(yintercept = 1, color="darkblue", size = 1) +
  xlab("Date") + ylab("Percent Vaccinated") + 
  theme_few() + scale_y_continuous(breaks = seq(0, 1, len = 11), labels=percent) + 
  theme(panel.grid.major.y = element_line(colour="lightgrey", size = .25)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  ggtitle("Percentage of the US Population Vaccinated - Without J&J")
population_vaccinated_nojj

## -- FIGURE 5 - STOCK OVER TIME - WITH J&J -- ##
# Create stock over time dataframe to create graph from
stock_withjj_base <- outcomes_table_withjj_base[, c("date", "stock_wk")] %>%
  rename(base = stock_wk)
stock_withjj_low <- outcomes_table_withjj_low[, c("date", "stock_wk")] %>%
  rename(low = stock_wk)
stock_withjj_high <- outcomes_table_withjj_high[, c("date", "stock_wk")] %>%
  rename(high = stock_wk)
stock_over_time_withjj <- merge(stock_withjj_low, stock_withjj_base)
stock_over_time_withjj <- merge(stock_over_time_withjj, stock_withjj_high)
stock_over_time_withjj <- stock_over_time_withjj %>%
  rename('Max 2.5 million/day' = low, 'Max 3 million/day' = base,
         'Max 3.5 million/day' = high) %>%
  select(date, "Max 2.5 million/day", "Max 3 million/day", "Max 3.5 million/day") %>%
  gather(key = "Scenario", value = "value", -date)

# Create graph
stock_graph_withjj <- ggplot(stock_over_time_withjj, aes(x=date, y=value)) +
  geom_line(aes(color = Scenario)) + 
  geom_point(aes(color=Scenario)) +
  scale_color_manual(values = c("darkred", "orange", "darkgreen")) +
  xlab("Date") + ylab("Total Stock") + 
  theme_few() + scale_y_continuous(labels=comma) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  theme(panel.grid.major.y = element_line(colour="lightgrey", size = .25)) +
  ggtitle("Total Stock of all Vaccines - Including J&J")
stock_graph_withjj

## -- FIGURE 6 - PCT POPULATION VACCINATION - WITH J&J -- ##
# Create percent vaccinated variables
outcomes_table_withjj_base <- outcomes_table_withjj_base %>%
  mutate(base = vaccinations_completed_tot/US_population)
outcomes_table_withjj_low <- outcomes_table_withjj_low %>%
  mutate(low = vaccinations_completed_tot/US_population)
outcomes_table_withjj_high <- outcomes_table_withjj_high %>%
  mutate(high = vaccinations_completed_tot/US_population)

# Create a new dataframe with percent vaccinated for graph
pct_vacc_table_withjj <- outcomes_table_withjj_base[,c("date", "base")]
pct_vacc_table_withjj <- cbind(pct_vacc_table_withjj, outcomes_table_withjj_low$low)
pct_vacc_table_withjj <- cbind(pct_vacc_table_withjj, outcomes_table_withjj_high$high)
pct_vacc_table_withjj <- pct_vacc_table_withjj %>%
  rename('Max 2.5 million/day' = 'outcomes_table_withjj_low$low', 
         'Max 3 million/day' = base,
         'Max 3.5 million/day' = 'outcomes_table_withjj_high$high') %>%
  select(date, "Max 2.5 million/day", "Max 3 million/day", "Max 3.5 million/day") %>%
  gather(key = "Scenario", value = "value", -date)

# Create graph
population_vaccinated_withjj <- ggplot(pct_vacc_table_withjj, aes(x=date, y=value)) +
  geom_line(aes(color = Scenario)) + 
  geom_point(aes(color=Scenario)) +
  scale_color_manual(values = c("darkred", "orange", "darkgreen")) +
  geom_hline(yintercept = 0.5, color="lightblue", size = 1) +
  geom_hline(yintercept = 0.75, color="blue", size = 1) +
  geom_hline(yintercept = 1, color="darkblue", size = 1) +
  xlab("Date") + ylab("Percent Vaccinated") + 
  theme_few() + scale_y_continuous(breaks = seq(0, 1, len = 11), labels=percent) + 
  theme(panel.grid.major.y = element_line(colour="lightgrey", size = .25)) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  ggtitle("Percentage of the US Population Vaccinated - Including J&J")
population_vaccinated_withjj
```